---
title: "R Notebook"
output: html_notebook
---


## Installing packages
```{r}

BiocManager::install("DESeq2")
install.packages("tidyverse")

```

## Loading libraries
```{r}
library("DESeq2")
library("variancePartition")
library("BiocParallel")
library("reshape2")
library("tidyverse")
```

```{r}
data_dir = file.path("~/Library/CloudStorage/GoogleDrive-dhaak@vt.edu/Shared drives/2022_SOSP_prelim/Data/")
figs_dir = file.path("~/Library/CloudStorage/GoogleDrive-dhaak@vt.edu/Shared drives/2022_SOSP_prelim/Figures/")
```


## Reading in the data
```{r}
bird_counts <- read.delim(file.path(data_dir, "bird_counts.txt"), comment.char="#", header = TRUE, row.names = 1)
bird_counts <- bird_counts[,-c(1:5)]
bird_counts <- bird_counts[rowSums(bird_counts)>=100,]
#bird_counts <- bird_counts[rowSums(bird_counts)<=50000,]
colnames(bird_counts)
countData <- bird_counts %>% 
  dplyr::rename(
    S11 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S11_S8_L001Aligned.sortedByCoord.out.bam, 
S12 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S12_S9_L001Aligned.sortedByCoord.out.bam, 
S13 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S13_S10_L001Aligned.sortedByCoord.out.bam,
S14 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S14_S11_L001Aligned.sortedByCoord.out.bam,
S16 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S16_S12_L001Aligned.sortedByCoord.out.bam,
S17 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S17_S13_L001Aligned.sortedByCoord.out.bam,
S18 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S18_S14_L001Aligned.sortedByCoord.out.bam,
S1 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S1_S1_L001Aligned.sortedByCoord.out.bam,
S22 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S22_S15_L001Aligned.sortedByCoord.out.bam,
S24 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S24_S16_L001Aligned.sortedByCoord.out.bam,
S25 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S25_S17_L001Aligned.sortedByCoord.out.bam,
S26 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S26_S18_L001Aligned.sortedByCoord.out.bam,
S27 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S27_S19_L001Aligned.sortedByCoord.out.bam,
S28 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S28_S20_L001Aligned.sortedByCoord.out.bam,
S2 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S2_S2_L001Aligned.sortedByCoord.out.bam,
S3 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S3_S3_L001Aligned.sortedByCoord.out.bam,
S4 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S4_S4_L001Aligned.sortedByCoord.out.bam,
S6 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S6_S5_L001Aligned.sortedByCoord.out.bam,
S7 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S7_S6_L001Aligned.sortedByCoord.out.bam,
S9 = X.projects.mcbee.haak.birdRNA_8387.02_mapped.02_1_zonoalign.8387.S9_S7_L001Aligned.sortedByCoord.out.bam)
View(countData)

colData <- read.table(file.path(data_dir, "Samples_files.txt"), header = TRUE, row.names = 1)
all(rownames(colData) %in% colnames(countData))
countData <- countData[, rownames(colData)]
all(rownames(colData) == colnames(countData)) 

```

```{r}
cor_samps<-round(cor(countData, method ="spearman"), digits = 2)
melted_cor_samps<-melt(cor_samps)
ggheatmap<-ggplot(data = melted_cor_samps, aes(x=Var1, y=Var2, fill = value)) + 
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
  theme(axis.text.y = element_text(size = 12))
ggheatmap<-ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "white", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank())
ggheatmap
```
```{r}
countDatadrops<-countData %>% 
  select(-S4, -S14)
cor_sampsdrops<-round(cor(countDatadrops, method ="spearman"), digits = 2)
melted_cor_samps<-melt(cor_sampsdrops)
ggheatmap<-ggplot(data = melted_cor_samps, aes(x=Var1, y=Var2, fill = value)) + 
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
  theme(axis.text.y = element_text(size = 12))
ggheatmap<-ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "white", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank())
ggheatmap

scale_drops<-scale(t(countDatadrops))
d <- dist(scale_drops)
hc<-hclust(d, method = "ward.D2")
plot(hc)
```


```{r}
dds<- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~ Site - 1)
dds<-DESeq(dds)
res01<-results(dds)
summary(res01)
names(res01)
mcols(res01)
reOrdered<-res01[order(res01$padj),]
resultsNames(dds)
rld<- rlog(dds, blind=TRUE)
vsd<-varianceStabilizingTransformation(dds, blind=TRUE)

select <- order(rowMeans(counts(dds, normalized=FALSE)), decreasing=TRUE)[1:100]
nt <- normTransform(dds) # defaults to log2(x+1)
log2.norm.counts <- assay(nt)[select,]

plotPCA(vsd, intgroup=c("Condition"))
```

```{r}
resultsNames(dds)
resNVTHER<-results(dds, contrast=c("Site","NVT","Heritage"), independentFiltering=TRUE, alpha=0.05, pAdjustMethod="BH", parallel=TRUE)
resSVTHER <- results(dds, contrast=c("Site", "SVT", "Heritage"), independentFiltering=TRUE, alpha=0.05, pAdjustMethod="BH", parallel=TRUE)

#install.packages("ashr")
require(ashr)

#Using LFC shrinkage to compare genes by L2FC
lfcNVTHER<- lfcShrink(dds, contrast=c("Site","NVT","Heritage"), type="ashr")
lfcSVTHER <- lfcShrink(dds, contrast=c("Site", "SVT", "Heritage") , type="ashr")
lfcRADHER <- lfcShrink(dds, contrast=c("Site", "Radford", "Heritage"), type = "ashr")
lfcNVTKEN<- lfcShrink(dds, contrast=c("Site","NVT","Kent"), type="ashr")
lfcSVTKEN <- lfcShrink(dds, contrast=c("Site", "SVT", "Kent") , type="ashr")
lfcRADKEN <- lfcShrink(dds, contrast=c("Site", "Radford", "Kent"), type = "ashr")
lfcNVTSTR<- lfcShrink(dds, contrast=c("Site","NVT","Stream"), type="ashr")
lfcSVTSTR <- lfcShrink(dds, contrast=c("Site", "SVT", "Stream") , type="ashr")
lfcRADSTR <- lfcShrink(dds, contrast=c("Site", "Radford", "Stream"), type = "ashr")

lfcNVTSVT<- lfcShrink(dds, contrast=c("Site","NVT","SVT"), type="ashr")
lfcSVTRAD <- lfcShrink(dds, contrast=c("Site", "SVT", "Radford") , type="ashr")
lfcRADNVT <- lfcShrink(dds, contrast=c("Site", "Radford", "NVT"), type = "ashr")
lfcNVTRAD<- lfcShrink(dds, contrast=c("Site","NVT","Radford"), type="ashr")
lfcKENHER<- lfcShrink(dds, contrast=c("Site","Kent","Heritage"), type="ashr")
lfcKENSTR<- lfcShrink(dds, contrast=c("Site","Kent","Stream"), type="ashr")
lfcSTRHER<- lfcShrink(dds, contrast=c("Site","Stream", "Heritage"), type="ashr")




lfcNVTHER = lfcNVTHER[order(lfcNVTHER$padj, na.last = NA), ]
lfcSVTHER = lfcSVTHER[order(lfcSVTHER$padj, na.last = NA), ]
lfcRADHER = lfcRADHER[order(lfcRADHER$padj, na.last = NA), ]
lfcNVTKEN = lfcKENHER[order(lfcKENHER$padj, na.last = NA), ]
lfcSVTKEN = lfcKENHER[order(lfcKENHER$padj, na.last = NA), ]
lfcRADKEN = lfcRADKEN[order(lfcRADKEN$padj, na.last = NA), ]
lfcNVTSTR = lfcNVTSTR[order(lfcNVTSTR$padj, na.last = NA), ]
lfcSVTSTR = lfcSVTSTR[order(lfcSVTSTR$padj, na.last = NA), ]
lfcRADSTR = lfcRADSTR[order(lfcRADSTR$padj, na.last = NA), ]
lfcNVTSVT = lfcNVTSVT[order(lfcNVTSVT$padj, na.last = NA), ]
lfcSVTRAD = lfcSVTRAD[order(lfcSVTRAD$padj, na.last = NA), ]
lfcRADNVT = lfcRADNVT[order(lfcRADNVT$padj, na.last = NA), ]
lfcNVTRAD = lfcNVTRAD[order(lfcNVTRAD$padj, na.last = NA), ]
lfcKENHER = lfcKENHER[order(lfcKENHER$padj, na.last = NA), ]
lfcKENSTR = lfcKENSTR[order(lfcKENSTR$padj, na.last = NA), ]
lfcSTRHER = lfcSTRHER[order(lfcSTRHER$padj, na.last = NA), ]

alpha = 0.01

```

# Contrasts with Heritage
```{r}
lfcNVTHER_tbl <- as_tibble(lfcNVTHER, rownames = "Gene_ID") %>% 
  mutate(group = "NVTHER") 
lfcSVTHER_tbl <- as_tibble(lfcSVTHER, rownames = "Gene_ID") %>% 
    mutate(group = "SVTHER")
lfcRADHER_tbl <- as_tibble(lfcRADHER, rownames = "Gene_ID") %>% 
    mutate(group = "RADHER")
lfcKENHER_tbl <- as_tibble(lfcKENHER , rownames = "Gene_ID") %>% 
  mutate(group = "KENHER" )
lfcSTRHER_tbl <- as_tibble(lfcSTRHER , rownames = "Gene_ID") %>% 
  mutate(group = "HERSTR" )

est_exp <- lfcNVTHER_tbl %>% 
  inner_join(lfcSVTHER_tbl, by = "Gene_ID") %>% 
  inner_join(lfcRADHER_tbl, by = "Gene_ID") %>% 
  inner_join(lfcKENHER_tbl, by = "Gene_ID") %>%
  inner_join(lfcSTRHER_tbl, by = "Gene_ID" ) %>% 
  select(-pvalue.x, -pvalue.y, -pvalue.x.x, -pvalue.y.y, -pvalue) %>% 
  dplyr::rename(
    L2FC_NVTHER = log2FoldChange.x,
    L2FC_SVTHER = log2FoldChange.y,
    L2FC_RADHER = log2FoldChange.x.x,
    L2FC_KENHER = log2FoldChange.y.y,
    L2FC_STRHER = log2FoldChange,
    padj_NVTHER = padj.x,
    padj_SVTHER = padj.y,
    padj_RADHER = padj.x.x,
    padj_KENHER = padj.y.y,
    padj_STRHER = padj
  ) %>% 
  mutate(
    sigUR = case_when(padj_NVTHER <= 0.05 & padj_SVTHER <= 0.05 ~ "UHER_only",
                           padj_KENHER <= 0.05 & padj_STRHER <= 0.05 ~ "RHER_only",
                           TRUE ~ "Unchaged")
  )
View(est_exp)

est_exp %>% 
  filter(sigUR == "UHER_only")
```

```{r}
lfcNVTKEN_tbl <- as_tibble(lfcNVTKEN, rownames = "Gene_ID") %>% 
  mutate(group = "NVTKEN") 
lfcSVTKEN_tbl <- as_tibble(lfcSVTKEN, rownames = "Gene_ID") %>% 
    mutate(group = "SVTKEN")
lfcRADKEN_tbl <- as_tibble(lfcRADKEN, rownames = "Gene_ID") %>% 
    mutate(group = "RADKEN")
lfcKENHER_tbl <- as_tibble(lfcKENHER , rownames = "Gene_ID") %>% 
  mutate(group = "KENHER" )
lfcKENSTR_tbl <- as_tibble(lfcKENSTR , rownames = "Gene_ID") %>% 
  mutate(group = "KENSTR" )

est_exp_KEN <- lfcNVTKEN_tbl %>% 
  inner_join(lfcSVTKEN_tbl, by = "Gene_ID") %>% 
  inner_join(lfcRADKEN_tbl, by = "Gene_ID") %>% 
  inner_join(lfcKENHER_tbl, by = "Gene_ID") %>%
  inner_join(lfcKENSTR_tbl, by = "Gene_ID" ) %>% 
  select(-pvalue.x, -pvalue.y, -pvalue.x.x, -pvalue.y.y, -pvalue) %>% 
  rename(
    L2FC_NVTKEN = log2FoldChange.x,
    L2FC_SVTKEN = log2FoldChange.y,
    L2FC_RADKEN = log2FoldChange.x.x,
    L2FC_KENHER = log2FoldChange.y.y,
    L2FC_KENSTR = log2FoldChange,
    padj_NVTKEN = padj.x,
    padj_SVTKEN = padj.y,
    padj_RADKEN = padj.x.x,
    padj_KENHER = padj.y.y,
    padj_KENSTR = padj
  ) %>% 
  mutate(
    sigUR = case_when(padj_NVTKEN <= 0.05 & padj_SVTKEN <= 0.05 ~ "UHER_only",
                           padj_KENHER <= 0.05 & padj_KENSTR <= 0.05 ~ "RHER_only",
                           TRUE ~ "Unchaged")
  )
View(est_exp_KEN)

est_exp_KEN %>% 
  filter(sigUR == "UHER_only" | sigUR == "RHER_only")
```

```{r}
lfcNVTSTR_tbl <- as_tibble(lfcNVTSTR, rownames = "Gene_ID") %>% 
  mutate(group = "NVTKEN") 
lfcSVTSTR_tbl <- as_tibble(lfcSVTKEN, rownames = "Gene_ID") %>% 
    mutate(group = "SVTKEN")
lfcRADSTR_tbl <- as_tibble(lfcRADKEN, rownames = "Gene_ID") %>% 
    mutate(group = "RADKEN")
lfcKENSTR_tbl <- as_tibble(lfcKENSTR , rownames = "Gene_ID") %>% 
  mutate(group = "KENHER" )
lfcSTRHER_tbl <- as_tibble(lfcSTRHER , rownames = "Gene_ID") %>% 
  mutate(group = "KENSTR" )

est_exp_STR <- lfcNVTSTR_tbl %>% 
  inner_join(lfcSVTSTR_tbl, by = "Gene_ID") %>% 
  inner_join(lfcRADSTR_tbl, by = "Gene_ID") %>% 
  inner_join(lfcKENSTR_tbl, by = "Gene_ID") %>%
  inner_join(lfcSTRHER_tbl, by = "Gene_ID" ) %>% 
  select(-pvalue.x, -pvalue.y, -pvalue.x.x, -pvalue.y.y, -pvalue) %>% 
  rename(
    L2FC_NVTSTR = log2FoldChange.x,
    L2FC_SVTSTR = log2FoldChange.y,
    L2FC_RADSTR = log2FoldChange.x.x,
    L2FC_KENSTR = log2FoldChange.y.y,
    L2FC_STRHER = log2FoldChange,
    padj_NVTSTR = padj.x,
    padj_SVTSTR = padj.y,
    padj_RADSTR = padj.x.x,
    padj_KENSTR = padj.y.y,
    padj_STRHER = padj
  ) %>% 
  mutate(
    sigUR = case_when(padj_NVTSTR <= 0.05 & padj_SVTSTR <= 0.05 ~ "UHER_only",
                           padj_KENSTR <= 0.05 & padj_STRHER <= 0.05 ~ "RHER_only",
                           TRUE ~ "Unchaged")
  )
View(est_exp_STR)

est_exp_STR %>% 
  filter(sigUR == "UHER_only" | sigUR == "RHER_only")
```

```{r}
# Some plotting
```{r reshape the count data}
normalized_counts = counts(dds, normalized=TRUE)
counts_reshape <- normalized_counts %>% 
  as_tibble(rownames = "Gene_ID") %>% 
  pivot_longer(cols = -Gene_ID, names_to = "SampleID", values_to = "Expression") %>% 
#  separate(SampleID, into = c("a", "b", "c", "c2", "Duke_ID", "Sample" ), sep = "_") %>% 
#  select(-a, -b , -c, -c2) %>% 
  view()
counts_reshape <- counts_reshape %>% 
 mutate(Habitat = recode(SampleID, S1 = "Rural",
                        S2 = "Rural",
                        S3 = "Urban",
                        S4 = "Urban",
                        S6 = "Rural",
                        S7 = "Urban",
                        S9 = "Urban",
                        S11 = "Rural",
                        S12 = "Urban",
                        S13 = "Rural",
                        S14 = "Rural",
                        S16 = "Rural",
                        S17 = "Rural",
                        S18 = "Urban",
                        S22 = "Urban",
                        S24 = "Rural",
                        S25 = "Urban",
                        S26 = "Rural",
                        S27 = "Urban",
                        S28 = "Rural",
                        )) %>% 
mutate(Site = recode(SampleID, S1 = "Heritage",
                        S2 = "Stream",
                        S3 = "NVT",
                        S4 = "SVT",
                        S6 = "Stream",
                        S7 = "SVT",
                        S9 = "SVT",
                        S11 = "Heritage",
                        S12 = "NVT",
                        S13 = "Stream",
                        S14 = "Kent",
                        S16 = "Kent",
                        S17 = "Stream",
                        S18 = "SVT",
                        S22 = "NVT",
                        S24 = "Heritage",
                        S25 = "Radford",
                        S26 = "Kent",
                        S27 = "Radford",
                        S28 = "Kent",
                        )) %>% 
  view()


counts_reshape %>% 
  ggplot(aes(SampleID, Expression, colors = Habitat)) +
  geom_boxplot() + 
  ylim(0,10000)
  theme_bw(16)
```


```{r}
library(pheatmap)
#find gene length normalized values 
#geneLengths <- counts$width
#rpk <- apply( subset(counts, select = c(-width)), 2, 
#              function(x) x/(geneLengths/1000))
#normalize by the sample size using rpk values
#tpm <- apply(rpk, 2, function(x) x / sum(as.numeric(x)) * 10^6)
colData2 <- colData %>% 
  select(-Brain_on_Ice, -Sacrifice_Time, -Playback_duration, -Seq_sample, -Date_collected, -Bird_ID)
selectedGenes <- names(sort(apply(normalized_counts, 1, var), 
                            decreasing = T)[1:100])
pheatmap(normalized_counts[selectedGenes,], 
             scale = 'row',
             annotation_col = colData2, 
             show_rownames = FALSE)
```
```{r}
set.seed(101)

scaleCounts <- scale(t(countData))
d = dist(scaleCounts)

hc=hclust(d, method="complete")
plot(hc)

# we have to transpore the matrix t()
# so that we calculate distances between samples
knormbird=kmeans(t(normalized_counts),centers=2)  

# number of data points in each cluster
table(knormbird$cluster)

library(cluster)

set.seed(101)
pamclu=cluster::pam(t(normalized_counts),k=4)
plot(silhouette(pamclu),main=NULL)

Ks=sapply(2:7,
    function(i) 
      summary(silhouette(pam(t(normalized_counts),k=i)))$avg.width)
plot(2:7,Ks,xlab="k",ylab="av. silhouette",type="b",
     pch=19)

library(NbClust)
nb = NbClust(data=t(normalized_counts), 
             distance = "euclidean", min.nc = 1,
        max.nc = 7, method = "kmeans",
        index=c("kl","ch","cindex","db","silhouette",
                "duda","pseudot2","beale","ratkowsky",
                "gap","gamma","mcclain","gplus",
                "tau","sdindex","sdbw"))
nb = NbClust(data=t(normalized_counts), 
             distance = "euclidean", min.nc = 1,
        max.nc = 3, method = "kmeans",
        index=c("gamma"))

table(nb$Best.nc[1,]) # consensus seems to be 3 clusters 
```


```{r}
 gene_plots <- function (x){
   plotsdf <- counts_reshape %>% 
     dplyr::filter(Gene_ID == x) %>%
     group_by(Habitat, Site) %>%
     dplyr::summarize(mean=mean(Expression), n = n(), se=sd(Expression)/sqrt(n))
   
   plot.x <- plotsdf %>%
     ggplot(aes(Habitat, mean, color = Site)) +
     stat_summary(fun="mean", geom="line", aes(group=factor(Habitat))) +
     geom_pointrange(aes(ymin = plotsdf$mean-plotsdf$se, ymax = plotsdf$mean + plotsdf$se)) +
     ylab(paste("Normalized mean Expression ", x, sep = " ")) + 
     theme_bw(16)
   
   return(plot.x)
 }

gene_plots("NPY")
gene_plots("NPY1R")
gene_plots("NPY2R")
gene_plots("NPY5R")
gene_plots("AVT4R")
gene_plots("AVT3R")
gene_plots("AVT")
gene_plots("AR")
gene_plots("ESR1")
gene_plots("ESR2")
gene_plots("ARO")
gene_plots("CRH")
gene_plots("CRHBP")
gene_plots("CRHR2")
gene_plots("NR3C1")
gene_plots("NR3C2")
gene_plots("LOC113459353") #Unknown
gene_plots("TH") #Tyrosine 
gene_plots("TRPC6") 
```

```{r}
BiocManager::install("WGCNA")
library("WGCNA")

normalized_counts <- assay(vsd) %>%
  t()

sft <- pickSoftThreshold(normalized_counts,
  dataIsExpr = TRUE,
  corFnc = cor,
  networkType = "signed"
)

sft_df <- data.frame(sft$fitIndices) %>%
  dplyr::mutate(model_fit = -sign(slope) * SFT.R.sq)
```

```{r}
ggplot(sft_df, aes(x = Power, y = model_fit, label = Power)) +
  # Plot the points
  geom_point() +
  # We'll put the Power labels slightly above the data points
  geom_text(nudge_y = 0.1) +
  # We will plot what WGCNA recommends as an R^2 cutoff
  geom_hline(yintercept = 0.80, col = "red") +
  # Just in case our values are low, we want to make sure we can still see the 0.80 level
  ylim(c(min(sft_df$model_fit), 1.05)) +
  # We can add more sensible labels for our axis
  xlab("Soft Threshold (power)") +
  ylab("Scale Free Topology Model Fit, signed R^2") +
  ggtitle("Scale independence") +
  # This adds some nicer aesthetics to our plot
  theme_classic()
```




```{r}
bwnet <- blockwiseModules(normalized_counts,
  maxBlockSize = 5000, # What size chunks (how many genes) the calculations should be run in
  TOMType = "signed", # topological overlap matrix
  power = 12, # soft threshold for network construction
  numericLabels = TRUE, # Let's use numbers instead of colors for module labels
  randomSeed = 1234, # there's some randomness associated with this calculation
  # so we should set a seed
)
```

```{r}
module_eigengenes <- bwnet$MEs

# Print out a preview
head(module_eigengenes)
```

```{r}
all.equal(colData$Condition, rownames(module_eigengenes))

des_mat <- model.matrix(~ colData$Condition)
# lmFit() needs a transposed version of the matrix
fit <- limma::lmFit(t(module_eigengenes), design = des_mat)

# Apply empirical Bayes to smooth standard errors
fit <- limma::eBayes(fit)

# Apply multiple testing correction and obtain stats
stats_df <- limma::topTable(fit, number = ncol(module_eigengenes)) %>%
  tibble::rownames_to_column("module")
View(stats_df)
```



```{r}
res.pca <- prcomp(bird_counts, scale = TRUE)
print(res.pca)
eig.val <- get_eigenvalue(res.pca)
eig.val

summary(bird_counts)

corrplot(var$cos2, is.corr=FALSE)
fviz_eig(res.pca, col.var="blue")

var <- get_pca_var(res.pca)
var

```

```{r}
library("corrplot")
corrplot(var$cos2, is.corr=FALSE)
fviz_pca_var(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("darkorchid4", "gold", "darkorange"),
             repel = TRUE
             )
```

```{r}
fviz_contrib(res.pca, choice = "ind", axes = 1:2)
```

## Trying subset version
```{r}
countDatasub <- countData %>% 
  select(S4,S7,S9,S11,S18,S24)
colDatasub <- colData %>% 
  filter( Site == "SVT" | Site == "Heritage")
```

```{r}
ddsSK<- DESeqDataSetFromMatrix(countData = countDatasub, colData = colDatasub, design = ~ Condition - 1)
ddsSK<-DESeq(ddsSK)
res01<-results(ddsSK)
summary(res01)
names(res01)
mcols(res01)
reOrdered<-res01[order(res01$padj),]
resultsNames(ddsSK)
rld<- rlog(ddsSK, blind=TRUE)
vsd<-varianceStabilizingTransformation(ddsSK, blind=TRUE)

select <- order(rowMeans(counts(ddsSK, normalized=FALSE)), decreasing=TRUE)[1:100]
nt <- normTransform(ddsSK) # defaults to log2(x+1)
log2.norm.counts <- assay(nt)[select,]

plotPCA(vsd, intgroup=c("Condition"))
```


```{r}
require(ashr)
ddsSK@colData$Condition <- as.factor(ddsSK@colData$Condition)
resultsNames(ddsSK)
lfcSK<- lfcShrink(ddsSK, contrast=c("Condition","Urban","Rural"), type="ashr")
lfcSK = lfcSK[order(lfcSK$padj, na.last = NA), ]
view(lfcSK)
```



# Some plotting
```{r reshape the count data}
normalized_countssub = counts(ddsSK, normalized=TRUE)
counts_reshapesub <- normalized_countssub %>% 
  as_tibble(rownames = "Gene_ID") %>% 
  pivot_longer(cols = -Gene_ID, names_to = "SampleID", values_to = "Expression") %>% 
#  separate(SampleID, into = c("a", "b", "c", "c2", "Duke_ID", "Sample" ), sep = "_") %>% 
#  select(-a, -b , -c, -c2) %>% 
  view()
counts_reshapesub <- counts_reshapesub %>% 
 mutate(Habitat = recode(SampleID, 
                        S4 = "Urban",
                        S7 = "Urban",
                        S9 = "Urban",
                        S11 = "Rural",
                        S18 = "Urban",
                        S24 = "Rural"
                        )) %>% 
mutate(Site = recode(SampleID, 
                        S4 = "SVT",
                        S7 = "SVT",
                        S9 = "SVT",
                        S11 = "Heritage",
                        S18 = "SVT",
                        S24 = "Heritage"
                        ))  
  view(counts_reshapesub)

lfcNVTSTR_subtbl <- as_tibble(lfcNVTSTR, rownames = "Gene_ID") %>% 
  mutate(group = "NVTSTR") 
view(lfcNVTSTR_subtbl)

lfcNVTSTR_subtbl %>%
  inner_join(counts_reshapesub, by = "Gene_ID")
  
  
  
```



```{r}
gene_plots <- function (x){
   plotsdf <- counts_reshapesub %>% 
     dplyr::filter(Gene_ID == x) %>%
     group_by(Habitat) %>%
     dplyr::summarize(mean=mean(Expression), n = n(), se=sd(Expression)/sqrt(n))
   
   plot.x <- plotsdf %>%
     ggplot(aes(Habitat, mean)) +
     stat_summary(fun="mean", geom="line", aes(group=factor(Habitat))) +
     geom_pointrange(aes(ymin = plotsdf$mean-plotsdf$se, ymax = plotsdf$mean + plotsdf$se)) +
     ylab(paste("Normalized mean Expression ", x, sep = " ")) + 
     theme_bw(16)
   
   return(plot.x)
 }

SVTHER_NPY<-gene_plots("NPY")
SVTHER_CNPY3<-gene_plots("CNPY3")
gene_plots("NPY1R")
gene_plots("NPY2R")
gene_plots("NPY5R")
gene_plots("AVT4R")
gene_plots("AVT3R")
gene_plots("AVT")
gene_plots("AR")
gene_plots("ESR1")
gene_plots("ESR2")
gene_plots("ARO")
gene_plots("CRH")
gene_plots("CRHBP")
gene_plots("CRHR2")
gene_plots("NR3C1")
gene_plots("NR3C2")
gene_plots("LOC113459353") #Unknown
gene_plots("TH") #Tyrosine 
gene_plots("TRPC6") 
gene_plots("BARHL1")#BarH like homeobox

ggsave(file.path(figs_dir, "SVT_Heritage_NPY.png"), SVTHER_NPY)
ggsave(file.path(figs_dir, "SVT_Heritage_CNPY3.png"), SVTHER_CNPY3)
```

```{r}
library(pheatmap)
#find gene length normalized values 
#geneLengths <- counts$width
#rpk <- apply( subset(counts, select = c(-width)), 2, 
#              function(x) x/(geneLengths/1000))
#normalize by the sample size using rpk values
#tpm <- apply(rpk, 2, function(x) x / sum(as.numeric(x)) * 10^6)
normalized_countssub<-counts(ddsSK, normalized=TRUE)
selectedGenes <- names(sort(apply(normalized_countssub, 1, var), 
                            decreasing = T)[1:100])
pheatmap(normalized_counts[selectedGenes,], 
             scale = 'row',
             annotation_col = colDatasub, 
             show_rownames = FALSE)
```